Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2021-09-15 vertical-thermal-review.Rmd
small edits

Sun leaves have anatomical, morphological, and physiological traits that reduce heat and water stress (Table `r table_leaf_traits`).
They are generally smaller, thicker, and more deeply lobed, resulting in higher $g_b$ and lower $T_{leaf}-T{air}$ [Fig. `r fig_leaf_T`, *REFS*]. 
Higher stomatal densities allow higher $g_s$.
Higher cuticle thickness and trichome density reduce water loss, and higher vein densities... [*REFS*].
Steeper leaf angles [@niinemetsAdjustmentFoliageStructure1998; @ballMaintenanceLeafTemperature1988] reduce radiation loads and thereby $T_{leaf}-T_{air}$.
*... are more pubescent, having higher vein density [@leighInfluenceLeafSize2017; @vogelSunLeavesShade1968a; @sackHowStrongIntracanopy2006; @zwienieckiHydraulicLimitationsImposed2004;@mathurPhotosyntheticEfficiencySun2018a], and a steeper leaf-angle [@niinemetsAdjustmentFoliageStructure1998; @ballMaintenanceLeafTemperature1988].* 
These traits collectively aid heat dissipation and regulate $T_{leaf}$-photosynthetic balance. 
In contrast, shade leaves have traits that maximize light capture in the diffuse-light environment with sunflecks (e.g., lower LMA), but their greater thermal capacitance makes them more prone to overheating than sun leaves [@leighInfluenceLeafSize2017; @casasSunShadeLeaves2011; @schymanskiStomatalControlLeaf2013b]. 
In open canopies, where light is comparatively homogeneous, exposed leaves are shaped more by $T_{air_{max}}$ and VPD stress than by light [@mediavillaFoliarPlasticityRelated2019a].

>>>>

Sun leaves have anatomical, morphological, and physiological traits that reduce heat and water stress (Table `r table_leaf_traits`). They are generally smaller, thicker, and more deeply lobed, resulting in higher $g_b$ and lower $T_{leaf}-T{air}$ [Fig. `r fig_leaf_T`, @leighInfluenceLeafSize2017; @vogelSunLeavesShade1968a; @sackHowStrongIntracanopy2006;@zwienieckiHydraulicLimitationsImposed2004;@mathurPhotosyntheticEfficiencySun2018a]. Higher stomatal and vein densities allow higher $g_s$ and facilitate higher transpiration respectively [@zwienieckiHydraulicLimitationsImposed2004]. Higher cuticle thickness and trichome density reduce water loss. Steeper leaf angles reduce radiation loads and thereby decrease $T_{leaf}-T{air}$ [@niinemetsAdjustmentFoliageStructure1998; @ballMaintenanceLeafTemperature1988]. These traits collectively aid heat dissipation and regulate T_leaf-photosynthetic balance. In contrast, shade leaves have traits that maximize light capture in the diffuse-light environment with sunflecks (e.g., lower LMA), but their greater thermal capacitance makes them more prone to overheating than sun leaves [@leighInfluenceLeafSize2017; @casasSunShadeLeaves2011; @schymanskiStomatalControlLeaf2013b]. In open canopies, where light is comparatively homogeneous, exposed leaves are shaped more by $T_{air_{max}}$ and VPD stress than by light [@mediavillaFoliarPlasticityRelated2019a].
